World Heritage Sites on Wikipedia: Controversies in a Context of Constrained Agency for Cultural Heritage Activism

### 2020-06-08 analysis/paper/paper.Rmd
fix typos in the paper

Only one of the controversial sites discussed above is in the group of articles where the talk page is longer than the article itself, and none of the site on the UNESCO heritage in danger list are in this group. The talk page for Mapungubwe is lengthy with expressions of concern about the accuracy of information in the article. In particular, editors were concerned about the depiction of Indigenous people, especially the degree to which early European colonists were aware of Indigenous communities, and the correct cultural affiliation of the Indigenous groups that originally occupied the site. These issues closely echo the themes identified by @meskell2011paris in her analysis of the WHC process. Although not notable for the length of its talk page, the discussions among editors of the article about Preah Vihear Temple include a strongly worded disagreement among seven editors about the legitimacy of the ownership claims by Thailand and Cambodia of the territory that includes the temple. The debate among Wikipedia editors mirrors very closely the political tensions surrounding the nomination of Preah Vihear. The talk page also has a much longer argument, mostly between two editors, about whether or not the Thai name of the site should be included in the text of the article. This debate also recapitiulates issues of which country has the stronger claim for ownership of the temple. Amid accusations of nationalism among the editors, one of them asks "Can we put aside politics and be more collaborative on the encyclopedia project?" and appeals to Wikipedia policies 'Third opinion' and 'Dispute resolution' to diffuse the tensions and refocus attention on the common goal. Beyond these two pages we found that the majority of content on CS-WHL article talk pages is about highly local topics, mostly article administration, editing processes, and quality control.

@@

Perhaps the most striking spatial pattern in our results is how much of the edits to CS-WHL articles in any part of the world come from the United States. The anonymous editors of articles about CS-WHL are generally not located in the same country as the site they are editing. This suggests a digital colonialism of world heritage information on Wikipedia, where heritage is interpreted and communicated by people who are not part of the descendant community traditionally associated with the site. This is an important observation because it is these editors that control the production of knowledge about CS-WHL articles, and the ability to control information about heritage has several implications. First is that the choices of these editors define community identity, that is, the community of Wikipedia editors, and also the broader community of Internet users that read Wikipedia, by indicating which sites are important and visible, and which are not by contributing to a digital 'authorized heritage discourse' [@smith2012constrained] that prioritizes their own self-interests. Second is that traditional communities associated with CS-WHL have little control over how their own heritage is represented on Wikipedia. We see this spatial pattern on Wikipedia as part of a trend of globalization of heritage, where places that had traditionally contributed to local or regional identities are turned into sanitized playgrounds for rich tourists [@bernbeck2004political]. Wikipedia may be seen as virtual playground for Americans to construct a universal heritage that reflects their neo-liberal cultural values, and validates their identities, perhaps reflecting a cultural rhetorcic 'heritage populism' [@reynie2016specter]. In this way, the representation of heritage on Wikipedia recapitulates a long-standing tradition of colonial subordination of one group's heritage and identity by a more powerful and dominant group.

>>>>

Only one of the controversial sites discussed above is in the group of articles where the talk page is longer than the article itself, and none of the site on the UNESCO heritage in danger list are in this group. The talk page for Mapungubwe is lengthy with expressions of concern about the accuracy of information in the article. In particular, editors were concerned about the depiction of Indigenous people, especially the degree to which early European colonists were aware of Indigenous communities, and the correct cultural affiliation of the Indigenous groups that originally occupied the site. These issues closely echo the themes identified by @meskell2011paris in her analysis of the WHC process. Although not notable for the length of its talk page, the discussions among editors of the article about Preah Vihear Temple include a strongly worded disagreement among seven editors about the legitimacy of the ownership claims by Thailand and Cambodia of the territory that includes the temple. The debate among Wikipedia editors mirrors very closely the political tensions surrounding the nomination of Preah Vihear. The talk page also has a much longer argument, mostly between two editors, about whether or not the Thai name of the site should be included in the text of the article. This debate also recapitulates issues of which country has the stronger claim for ownership of the temple. Amid accusations of nationalism among the editors, one of them asks "Can we put aside politics and be more collaborative on the encyclopedia project?" and appeals to Wikipedia policies 'Third opinion' and 'Dispute resolution' to diffuse the tensions and refocus attention on the common goal. Beyond these two pages we found that the majority of content on CS-WHL article talk pages is about highly local topics, mostly article administration, editing processes, and quality control.

@@

Perhaps the most striking spatial pattern in our results is how much of the edits to CS-WHL articles in any part of the world come from the United States. The anonymous editors of articles about CS-WHL are generally not located in the same country as the site they are editing. This suggests a digital colonialism of world heritage information on Wikipedia, where heritage is interpreted and communicated by people who are not part of the descendant community traditionally associated with the site. This is an important observation because it is these editors that control the production of knowledge about CS-WHL articles, and the ability to control information about heritage has several implications. First is that the choices of these editors define community identity, that is, the community of Wikipedia editors, and also the broader community of Internet users that read Wikipedia, by indicating which sites are important and visible, and which are not by contributing to a digital 'authorized heritage discourse' [@smith2012constrained] that prioritizes their own self-interests. Second is that traditional communities associated with CS-WHL have little control over how their own heritage is represented on Wikipedia. We see this spatial pattern on Wikipedia as part of a trend of globalization of heritage, where places that had traditionally contributed to local or regional identities are turned into sanitized playgrounds for rich tourists [@bernbeck2004political]. Wikipedia may be seen as virtual playground for Americans to construct a universal heritage that reflects their neo-liberal cultural values, and validates their identities, perhaps reflecting a cultural rhetoric of 'heritage populism' [@reynie2016specter]. In this way, the representation of heritage on Wikipedia recapitulates a long-standing tradition of colonial subordination of one group's heritage and identity by a more powerful and dominant group.
